Writing likelihood functions in R
========================================================

I'm having trouble with the syntax for writing likelihood functions in R. This is important since we want to create lots of dummy examples of small phylogenies, run them both on diversitree and any parallel approach we come up with.

Creating own function, let's start simple
Let’s take a look with some data found in ?mle2

```{r}
x = 0:10
y = c(26, 17, 13, 12, 20, 5, 9, 8, 5, 4, 8)
d = data.frame(x, y)
```

Here, we’re fitting a possion model that depends on an intercept term plus a linear term. The exp() is mainly there to make sure that the value of lambda doesn’t go negative, which isn’t allowed (it would imply a negative number of occurrences for our outcome of interest).

```{r}
require(bbmle)
#using R's native optim function()

fit0 = mle2(y ~ dpois(lambda= exp(intercept + slope * x)), start = list(intercept = mean(y), slope = 0), data = d)


simpleF <- function(slope,intercept) {
  -sum(dpois(y,  lambda=exp(intercept+slope*x), log = TRUE))}

fit1 = mle2(simpleF, start = list( intercept= mean(y), slope = 0))

fit0

fit1
```

We want to write a function that looks like simpleF, not as the native mle2 ~ style. 

Questions
----------

In our analogy, what will our x's and y's be? 

How many parameters will we need to fit?

